REVISITING SEMI-CONTINUOUS HIDDEN MARKOV MODELS

### 2011-09-26 misc/papers/icassp12_tied/icassp12.tex
updates to semicont paper
git-svn-id: https://svn.code.sf.net/p/kaldi/code/sandbox/discrim@514 5e6a8d80-dfce-4ca6-a32a-6e07a63d50c8

%
% Single address.
% ---------------
\name{
K.~Riedhammer$^1$,
T.~Bocklet$^1$,
A.~Ghoshal$^2$,
D.~Povey$^3$
}
\address{
$^1$Pattern Recognition Lab, University of Erlangen-Nuremberg, \sc{Germany}\\
$^2$Spoken Language Systems, Saarland University, \sc{Germany}\\
$^3$Microsoft Research, Redmond, WA, {\sc USA}\\
{\small\tt korbinian.riedhammer@informatik.uni-erlangen.de}
}

@@

\begin{abstract}
In the past decade, semi-continuous hidden Markov models (HMMs) have not attracted 
much attention in the speech recognition community. The strongly growing amount
of training data and continuous improvements of the initialization and training
procedures led to the impression that continuous HMMs are the best choice of
acoustic model.
%
However, recent work on tasks involving cross-lingual speech recognition and
languages with little training data faced the same the old problem -- millions
of parameters require a large amount of (unavailable) training data -- thus people are
investigating methods of reducing the number of parameters while maintaining or
extending the modeling capabilities of continuous models.
%
In this work, we compare continuous, semi-continuous, two-level tree based 
multiple-codebook semi-continuous with full covariance matrices and subspace Gaussian
mixture models.
%
Experiments using on the RM and WSJ corpora show that a semi-continuous system
can still yield competitive results while using fewer Gaussian components.
\end{abstract}

@@

\section{Introduction}
\label{sec:intro}
In the past years, semi-continuous hidden Markov models (SC-HMMs) \cite{huang1989shm}
have not attracted much attention in the automatic speech recognition (ASR) community. 
Most major frameworks (e.g.~CMU {\sc Sphinx} or HTK) more or less retired
semi-continuous models in favor of continuous models, which performed 
considerably better with the growing amount of training data available.
%
However, topics like cross-lingual model initialization and model training
for languages with little training data became more and more popular recently,
along with efforts on how to reduce the number of parameters of a continuous
system while maintaining or even extending the modeling capabilities and thus
the robustness and performance.
%
Typical techniques to reduce the number of parameters for continuous
systems include generalizations of the phonetic context modeling to reduce the 
number of states and state-tying to reduce the number of Gaussians.

It is well known that context dependent states significantly improve the
recognition performance (e.g.~from monophone to triphone). However for a
continuous system, increasing the number of states not only requires more
Gaussians to be estimated, but also implies that the training data for each
state is reduced.
%
This is the probably strongest advantage of semi-continuous models: While the
codebook Gaussians are always estimated on the whole data set, the only state
specific variables are the weights of the Gaussians, i.e.~increasing the number
of states only slightly increases the number of parameters, and the Gaussian
means and variances are still estimated using all the data; the reduced number
of Gaussian parameters for a large data set also allows to robustly estimate
full covariances.
%
This allows to reliably estimate arbitrary large phonetic context 
(``polyphones''), that showed good performance for the recognition of 
spontaneous  speech \cite{schukattalamazzini1994srf,schukattalamazzini1995as} 
while having a rather small number of parameters.
%
Further advantages of the shared codebook Gaussians are that they need to be 
evaluated only once per frame, and that the codebook can be initialized, trained
and adapted in a solely acoustic way.

@@

\subsection{Continuous Models}
For continuous models, every leaf of the tree is assigned an separate
Gaussian mixture model. We initialize the mixtures with a single component
each, and subsequently allocate more components by splitting components every 
iteration until a certain total number of components is reached (in this case
about 9000). We allocate the Gaussians proportional to a small power (e.g.~0.2) 
of the state's occupation count.
Formally, the emission probability of tree leaf $j$ is computed as
\begin{equation}
p(\x | j) = \sum_{i=1}^{N_j} c_{ji} \nv(\x; \m_{ji}, \k_{ji}) 
\end{equation}
where $N_j$ is the number of Gaussians assigned to $s_j$, and the $\m_{ji}$ and
$\k_{ji}$ are the means and covariance matrices of the mixtures.

@@

\section{Summary}
In this article, we evaluated semi-continuous and two-level tree based
semi-continuous acoustic models of the {\sc Kaldi} toolkit in the context of
both continuous and SGMM models.
%
While the first could not produce convincing results, the two-level tree 
based multiple-codebook architecture shows promising performance, especially for 
little training data.

>>>>

%
% Single address.
% ---------------
\name{
K.~Riedhammer$^1$,
T.~Bocklet$^1$,
A.~Ghoshal$^{2,3}$,
D.~Povey$^4$
}
\address{
$^1$Pattern Recognition Lab, University of Erlangen-Nuremberg, \sc{Germany}\\
$^2$Spoken Language Systems, Saarland University, \sc{Germany}\\
$^3$Centre for Speech Technology Research, University of Edinburgh, UK\\
$^4$Microsoft Research, Redmond, WA, {\sc USA}\\
{\small\tt korbinian.riedhammer@informatik.uni-erlangen.de}
}

@@

\begin{abstract}
In the past decade, semi-continuous hidden Markov models (SC-HMMs) have not 
attracted much attention in the speech recognition community. Growing amounts 
of training data and increasing sophistication of model estimation led to the 
impression that continuous HMMs are the best choice of acoustic model.
%
However, recent work on recognition of under-resourced languages faces the same the old problem of estimating a large number of parameters from limited amounts 
of labeled speech.
This has led to a renewed interest in methods of reducing the number of parameters while maintaining or 
extending the modeling capabilities of continuous models.
%
In this work, we compare continuous, semi-continuous, %two-level tree based 
multiple-codebook semi-continuous with full covariance matrices and subspace Gaussian
mixture models.
%
Experiments using on the RM and WSJ corpora show that a semi-continuous system
can still yield competitive results while using fewer Gaussian components.
\end{abstract}

@@

\section{Introduction}
\label{sec:intro}
In the past years, semi-continuous hidden Markov models (SC-HMMs) \cite{huang1989shm}
have not attracted much attention in the automatic speech recognition (ASR) community. 
Most major frameworks (e.g.~CMU {\sc Sphinx} or HTK) more or less retired
semi-continuous models in favor of continuous models, which performed 
considerably better with the growing amount of available training data.
%
However, topics like cross-lingual model initialization and model training
for languages with little training data became more and more popular recently,
along with efforts on how to reduce the number of parameters of a continuous
system while maintaining or even extending the modeling capabilities and thus
the robustness and performance.
%
Typical techniques to reduce the number of parameters for continuous
systems include generalizations of the phonetic context modeling to reduce the 
number of states and state-tying to reduce the number of Gaussians.

It is well known that context dependent states significantly improve the
recognition performance (e.g.~from monophone to triphone). However for a
continuous system, increasing the number of states not only requires more
Gaussians to be estimated, but also implies that the training data for each
state is reduced.
%
This is where semi-continuous models hold an advanatge: the only state specific 
variables are the weights of the Gaussians, while the codebook Gaussians 
themselves are always estimated on the whole data set. In other words, increasing the number of states leads to only modest increase in the total number of 
parameters. The Gaussian means and variances are estimated using all of the available data, making it possible to robustly estimate even full covariance matrices.
%
This also allows to reliably estimate arbitrarily large phonetic context 
(``polyphones''), which has been shown to achieve good performance for the 
recognition of spontaneous  speech \cite{schukattalamazzini1994srf,schukattalamazzini1995as} 
while having a relatively small number of parameters.
%
Further advantages of the shared codebook Gaussians are that they need to be 
evaluated only once per frame, and that the codebook can be initialized, trained
and adapted using even untranscribed speech.

@@

\subsection{Continuous Models}
For continuous models, every leaf of the tree is assigned an separate
Gaussian mixture model. Formally, the emission probability of tree leaf $j$ is 
computed as
\begin{equation}
p(\x | j) = \sum_{i=1}^{N_j} c_{ji} \nv(\x; \m_{ji}, \k_{ji}) 
\end{equation}
where $N_j$ is the number of Gaussians assigned to $s_j$, and the $\m_{ji}$ and
$\k_{ji}$ are the means and covariance matrices of the mixtures.

We initialize the mixtures with a single component each, and subsequently 
allocate more components by splitting components at every iteration until a 
pre-determined number of components is reached (in this case 9000). We allocate 
the Gaussians proportional to a small power (e.g.~0.2) of the state's 
occupation count.

@@

\section{Summary}
In this article, we compared two types of semi-continuous hidden Markov models, 
one using a single codebook and the other using multiple codebooks based 
on a two-level phonetic decision tree, with both continuous HMMs and SGMMs.
%
While the first could not produce convincing results, the two-level tree 
based multiple-codebook architecture shows promising performance, especially 
with limited training data.
